INTRODUCTION
    When is Novelty used?
    What is deception?
    How is our problem deceptive?
    Works done with Novelty search, deceptive problems, antenna...?

METHODS
Overview
    aed and genscreen recent crystallization methods
    chemical space unexplored
    considered an enhanced version



OTHER
Todo
    chemicals not in hwi-compounds
    initial population novelty how??
    naming combine link with midusha



LATER
Results
    overlap excel file, and other tables
    

https://ai.stackexchange.com/questions/240/what-exactly-are-genetic-algorithms-and-what-sort-of-problems-are-they-good-for/244#244
https://ai.stackexchange.com/questions/1296/how-does-novelty-search-work

In contrast, I believe the thinking is that novelty affords a kind of dynamic fitness, tending to push the population away from previously discovered areas.

However, despite their widespread use for the purpose, note that GAs are actually not function optimizers - GA mechanisms tend not to explore 'outlying' regions of the search space in the hope of finding some distant high quality solution, but rather to cluster around more easily attainable peaks in the 'fitness landscape'.


This perspective leads to a key idea in this dissertation that approaches the problems in both EC and artificial life in a new way: Instead of modeling natural evolution with the hope that novel individuals will be continually discovered, it is possible to create an openended dynamic by simply searching directly for novelty. Thus this dissertation introduces the novelty search algorithm, which searches with no objective; rather, it continually searches for novel behaviors in the search space. That is, the algorithm searches explicitly for behaviors not demonstrated by prior individuals.

Furthermore, because novelty search indirectly rewards lineages better able to produce novelty, it may also encourage evolvability, or the ability to further evolve, which is a desirable property that is apparent in natural evolution but rarely in EC. In contrast to novelty search, the pressure to optimize an objective can discourage evolvability when it converges upon a local optimum


Include contributions section:
    the research hypothesis section
    
    


Include Outline section. Section 1 describes and stuffs in detail


definition:
cocktails - equivalent to individual


explain runtime optimization in detail in methods, and again write in short in experiments section


DONEEE write equations of genetic algorithm



IMPORTANTTTT these are the cocktails generated by NS that were never found in 100 runs of GA.
PRELOADING ALL FEATURE VECTORS OF ALL COCKTAILS



hits excel sheets
screenshot of the program?

family = triad of ph, salt, precipitant
check if correctly implemented in program




goal is to explore search space effectively.... search for algorithms, found novelty search, like nature, read eplex page, 